Global code motion of congruent instructions

### 2017-04-18 paper.tex
new page to put references

\section{Introduction}
\label{sec:intro}
Compiler techniques to remove redundant computations are composed of an analysis
phase that detects identical computations in the program and a transformation
phase that reduces the number of run-time computations.  Classical scalar
optimizations like CSE \cite{dragonbook} work very well on single basic-blocks but
when it comes to detect redundancies across basic-blocks these techniques fall
short: more complex passes like GCSE and \PRE{} have been designed to handle these
cases based on dataflow analysis \cite{morel1979global}.  At first these
techniques were described in the classical data-flow analysis framework, and
later the use of the \SSA{} representation lowered the cost in terms of compilation
time \cite{briggs1994effective,chow1997new,kennedy1999partial} and brought these
techniques in the main stream: nowadays \SSA{} based \PRE{} is available in every
industrial compiler.

@@

\citet{dhamdhere1988fast}, \citet{steven1997advanced} mention
code hoisting in a data flow framework. A list of Very Busy Expressions (VBE)
computed which are hoisted in a basic-block where the expression is
anticipable (all the operands are available). This algorithm would hoist as far
as possible without regarding the impact on register pressure and as such a cost
model will be required.  Also the description of VBE is based on the classic
dataflow model and an adaptation to a sparse \SSA{} representation is required.

@@

\citet{shobaki2013} also recently presented a combinatorial global
scheduler with reasonable performance improvements. It is possible that both
Barany and Shobaki's implmentation will have similar results when compiled for
same target architecture. Also, both suffer from the same problem, although
Shobaki not so much, of large compile times which is not feasible for industrial
compilers like \GCC{} and \LLVM{}. With the algorithm described in the next
section, we got a reduction in register spills, improved inliner heuristics,
improved compile time and some performance improvements on \SPEC{} benchmarks.

@@

\begin{verbatim}

// Return true if the merge-set of source node of a visited J-edge (incoming edge of
// lnode) is not the subset of the merge-set of lnode.
bool still_inconsistent(Node lnode, JEdges JE, Visited V, MergeSet MergeSet)
  for e in (all incoming edges to lnode)
    if (e is in JE && e in V)
      if (MergeSet(Source Node of e) is not a subset of MergeSet(lnode))
        return true
  return false

@@

// Construct merge-set of each node in Control-Flow-Graph G of a function.
void  constructMergeSet(CFG G) {
  B = Breadth First Order of G
  JE = JEdges of G
  DL = List of Path length (from root) of each node in G.

  do // Call until a fixed point is reached.
    Repeat = constructMergeSet_1(B, JE, DL)
  while (Repeat)
}
\end{verbatim}

@@

\section{Conclusion and Future Work}
\label{sec:future-work}

@@

>>>>

\section{Introduction}
\label{sec:intro}
Compiler techniques to remove redundant computations are composed of an analysis
phase that detects identical computations in the program and a transformation
phase that reduces the number of run-time computations.  Classical scalar
optimizations like CSE \cite{dragonbook} work very well on single basic-blocks but
when it comes to detect redundancies across basic-blocks these techniques fall
short: more complex passes like GCSE and \PRE{} have been designed to handle these
cases based on data-flow analysis \cite{morel1979global}.  At first these
techniques were described in the classical data-flow analysis framework, and
later the use of the \SSA{} representation lowered the cost in terms of compilation
time \cite{briggs1994effective,chow1997new,kennedy1999partial} and brought these
techniques in the main stream: nowadays \SSA{} based \PRE{} is available in every
industrial compiler.

@@

\citet{dhamdhere1988fast} and \citet{steven1997advanced} mention
code hoisting in a data-flow framework. A list of Very Busy Expressions (VBE)
computed which are hoisted in a basic-block where the expression is
anticipable (all the operands are available). This algorithm would hoist as far
as possible without regarding the impact on register pressure and as such a cost
model will be required.  Also the description of VBE is based on the classic
data-flow model and an adaptation to a sparse \SSA{} representation is required.

@@

\citet{shobaki2013} also recently presented a combinatorial global
scheduler with reasonable performance improvements. It is possible that both
Barany and Shobaki's implmentation will have similar results when compiled for
same target architecture. Also, both suffer from the same problem, although
Shobaki not so much, of large compile times which would not suit in industrial
compilers like \GCC{} and \LLVM{}. With the algorithm described in the next
section, we got a reduction in register spills, improved inliner heuristics,
improved compile time, and show performance improvements on \SPEC{} benchmarks.

@@

\begin{verbatim}
// Return true if the merge-set of source node of a visited J-edge (incoming edge of
// lnode) is not the subset of the merge-set of lnode.
bool still_inconsistent(Node lnode, JEdges JE, Visited V, MergeSet MergeSet)
  for e in (all incoming edges to lnode)
    if (e is in JE && e in V)
      if (MergeSet(Source Node of e) is not a subset of MergeSet(lnode))
        return true
  return false

@@

// Construct merge-set of each node in Control-Flow-Graph G of a function.
void constructMergeSet(CFG G)
  B = Breadth First Order of G
  JE = JEdges of G
  DL = List of Path length (from root) of each node in G.
  do // Call until a fixed point is reached.
    Repeat = constructMergeSet_1(B, JE, DL)
  while (Repeat)
\end{verbatim}

@@

\section{Conclusions and Future Work}
\label{sec:future-work}

@@

\newpage
